Variational Monte Carlo on Bosonic Systems

### 2018-04-05 project_1/article/article.tex
Spell check section on MC integration

    \subsection{Monte Carlo integration}
        Unfortunately the evaluation of the expectation value of the energy for
        a given trial wavefunction involves a large integral.
        \begin{align}
            E
            &= \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
            =
            \frac{\int\dd\vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
            {\int\dd\vf{r}|\Psi_T|^2},
        \end{align}
        where $\vf{r} = (\vf{r}_1, \vf{r}_2, \dots \vf{r}_N)$ are the
        positions\footnote{This can include both spatial and spin parts, but we
        will in the remainder of the project ignore spin.} of the particles
        contained in the system.

        To get around\footnote{Not exactly \emph{get around} per se, but make
        the problem solvable at all.} the trouble of evaluating the
        multidimensional-integral we employ the method of Monte Carlo
        integration. This consists of approximating an integral of the
        expectation value of a certain quantitity by a sum.
        \begin{align}
            \expval{x}
            &=
            \int_{\mathbb{R}}xp(x)\dd x
            \approx
            \frac{1}{N}\sum_{i = 1}^N x_ip(x_i),
        \end{align}
        where $x$ is a random variable with a probability density function
        $p(x)$ and $\{x_i\}$ for $i \in \{1, \dots, N\}$ is the set of drawn
        random variables. Here $N$ is the number of Monte Carlo samples chosen
        to approximate the integral. By the law of large numbers we expect that
        \begin{align}
            \lim_{N \to \infty}\frac{1}{N}\sum_{i = 1}^N x_ip(x_i)
            = \expval{x}.
        \end{align}

        For us to be able to take Monte Carlo integration to the quantum realm
        we have to construct a probability density function from the
        wavefunction. Luckily the wavefunction squared is just such a quantity.
        Given a normalized wavefunction $\ket{\psi}$ we can compute the
        expectation value of a quantum observable $\mathcal{O}$ using Monte
        Carlo integration.
        \begin{align}
            \expval{\mathcal{O}}
            &= \bra{\psi}\mathcal{O}\ket{\psi}
            =
            \int\dd\vf{r}
            \psi^{*}(\vf{r})\mathcal{O}\psi(\vf{r})
            \\
            &=
            \int\dd\vf{r}
            |\psi(\vf{r})|^2
            \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})},
        \end{align}
        where we have multiplied and divided by the wavefunction. We now define
        the probability density function, $p(\vf{r})$, as
        \begin{align}
            p(\vf{r}) = |\psi(\vf{r})|^2,
        \end{align}
        and the local quantum observable
        \begin{align}
            \mathcal{O}_L(\vf{r})
            &= \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
        \end{align}
        We see that if the chosen wavefunction is an eigenstate of the quantum
        operator, the local quantum observable becomes a constant.  We can thus
        approximate the expecatation value of the quantum observable using Monte
        Carlo integration.
        \begin{align}
            \expval{O}
            &=
            \int\dd\vf{r}
            p(\vf{r})
            \mathcal{O}_L(\vf{r})
            \approx
            \frac{1}{N}\sum_{i = 1}^N
            p(\vf{r}_i)\mathcal{O}_L(\vf{r}_i).
        \end{align}
        In the case of a non-normalized wavefunction we need to divide by a
        normalization factor in the probability density function.

>>>>

    \subsection{Monte Carlo integration}
        Unfortunately the evaluation of the expectation value of the energy for
        a given trial wavefunction involves a large integral.
        \begin{align}
            E
            &= \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
            =
            \frac{\int\dd\vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
            {\int\dd\vf{r}|\Psi_T|^2},
        \end{align}
        where $\vf{r} = (\vf{r}_1, \vf{r}_2, \dots \vf{r}_N)$ are the
        positions\footnote{This can include both spatial and spin parts, but we
        will in the remainder of this project ignore spin.} of the particles
        contained in the system.

        To get around\footnote{Not exactly \emph{get around} per se. We still
        have to evaluate the integral, but we make the problem solvable at all.}
        the trouble of evaluating the multidimensional-integral we employ the
        method of Monte Carlo integration. This consists of approximating an
        integral of the expectation value of a certain quantitity by a sum.
        \begin{align}
            \expval{x}
            &=
            \int_{\mathbb{R}}xp(x)\dd x
            \approx
            \frac{1}{N}\sum_{i = 1}^N x_ip(x_i),
        \end{align}
        where $x$ is a random variable with a probability density function
        $p(x)$ and $\{x_i\}$ for $i \in \{1, \dots, N\}$ is the set of drawn
        random variables. Here $N$ is the number of Monte Carlo samples chosen
        to approximate the integral. By the law of large numbers we expect that
        \begin{align}
            \lim_{N \to \infty}\frac{1}{N}\sum_{i = 1}^N x_ip(x_i)
            = \expval{x}.
        \end{align}

        For us to be able to take Monte Carlo integration to the quantum realm
        we have to construct a probability density function from the
        wavefunction. Luckily the wavefunction squared is just such a quantity.
        Given a normalized wavefunction $\ket{\psi}$ we can compute the
        expectation value of a quantum operator $\mathcal{O}$ using Monte
        Carlo integration.
        \begin{align}
            \expval{\mathcal{O}}
            &= \bra{\psi}\mathcal{O}\ket{\psi}
            =
            \int\dd\vf{r}
            \psi^{*}(\vf{r})\mathcal{O}\psi(\vf{r})
            \\
            &=
            \int\dd\vf{r}
            |\psi(\vf{r})|^2
            \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})},
        \end{align}
        where we have multiplied and divided by the wavefunction. We now define
        the probability density function, $p(\vf{r})$, as
        \begin{align}
            p(\vf{r}) \equiv |\psi(\vf{r})|^2,
        \end{align}
        and the local quantum operator
        \begin{align}
            \mathcal{O}_L(\vf{r})
            &\equiv \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
        \end{align}
        We see that if the chosen wavefunction is an eigenstate of the quantum
        operator, the local quantum operator becomes a constant.  We can thus
        approximate the expecatation value of the quantum operator using Monte
        Carlo integration.
        \begin{align}
            \expval{O}
            &=
            \int\dd\vf{r}
            p(\vf{r})
            \mathcal{O}_L(\vf{r})
            \approx
            \frac{1}{N}\sum_{i = 1}^N
            p(\vf{r}_i)\mathcal{O}_L(\vf{r}_i).
        \end{align}
        In the case of a non-normalized wavefunction we need to divide by a
        normalization factor in the probability density function.
